---
title: "Simulation of the DSK SFC model sample with LHS"
author: "Mario Martinoli"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

# DSK SFC model simulation

This R Markdown simulates the DSK SFC model for different CoPs and MC runs. Use the CSV as a design table, but do not pass it directly to the model. The model accepts one JSON parameter file and one random seed per execution. The workflow is: CSV row --> modified JSON --> 1,000 seeded model runs --> extract 8 series --> save one 8 × 600 × 1,000 array.
Repeating this for 1,000 CSV rows produces 1,000 arrays—or equivalently a conceptual four-dimensional object with dimensions 8 × 600 × 1000 × 1000.

## Generate the parameter design

```{r parameters}
library(lhs)

set.seed(1)

bounds <- data.frame(
  parameter = c(
    "Gamma", "chi", "psi1", "psi3", "alfa",
    "taylor1", "taylor2", "taylor", "kappa"
  ),
  lower = c(0.05, -1.50, 0.10, 0.10, 0.10, 1.00, 0.00, 0.50, 0.50),
  upper = c(0.25, -1.25, 0.50, 0.50, 0.50, 1.50, 0.50, 0.95, 0.95)
)

u <- randomLHS(1000, nrow(bounds))

Sample <- sweep(u, 2, bounds$upper - bounds$lower, "*")
Sample <- sweep(Sample, 2, bounds$lower, "+")
colnames(Sample) <- bounds$parameter

stopifnot(anyDuplicated(as.data.frame(Sample)) == 0L)

dir.create("calibration", recursive = TRUE, showWarnings = FALSE)
write.csv(
  Sample,
  "calibration/Sample1.csv",
  row.names = FALSE,
  quote = FALSE
)
```

## Compile the model

The executable accepts a JSON file, run name, seed, and output flags as documented in the repository’s command-line instructions (https://github.com/CoMoS-SA/Reissl_2025/blob/main/README.md#command-line-arguments).

Run in your terminal:

git clone https://github.com/CoMoS-SA/Reissl_2025.git
cd Reissl_2025

cmake -S . -B build
cmake --build build -j 8

## Use the correct output columns

The normal results file contains 83 whitespace-separated columns and no header. The required mapping is:

| Array variable | Model expression | Output column |
|---|---|---:|
| GDP | `GDP_r(1)` | 2 |
| Consumption | `Consumption_r` | 3 |
| Investment | `Investment_r` | 4 |
| Employment rate | `1-U(1)` | 5 |
| CPI level | `cpi(1)` | 34 |
| CB lending/policy rate | `r` | 50 |
| Emissions | `Emiss1_TOT+Emiss2_TOT+Emiss_en` | 7 |
| Energy use | `D_en_TOT` | 8 |

These positions come directly from SAVE() in dsk_sfc_main.cpp (https://github.com/CoMoS-SA/Reissl_2025/blob/main/dsk_sfc_main.cpp#L6685-L6788).
Important: column 6 is not the CPI level. It is cpi(1)/cpi(5), the four-period gross inflation factor. If that is what you intended by “CPI,” use column 6 instead of 34. Similarly, column 50 is the central-bank monetary-policy rate; the average commercial-bank loan rate is column 82.

## R driver for one CoP

Save this as run_one_cop.R. Each invocation handles one CSV row and creates one array.

```{r compiler}
library(jsonlite)

args <- commandArgs(trailingOnly = TRUE)

if (length(args) < 5L) {
  stop(
    "Usage: Rscript run_one_cop.R COP_ID DESIGN.csv ",
    "BASE.json DSK_EXECUTABLE RESULT_DIR [N_MC]"
  )
}

cop_id      <- as.integer(args[1])
design_path <- normalizePath(args[2], mustWork = TRUE)
base_json   <- normalizePath(args[3], mustWork = TRUE)
source_exe  <- normalizePath(args[4], mustWork = TRUE)
result_dir  <- args[5]
n_mc        <- if (length(args) >= 6L) as.integer(args[6]) else 1000L

n_steps <- 600L
n_model_columns <- 83L

output_columns <- c(
  GDP             = 2L,
  Consumption     = 3L,
  Investment      = 4L,
  Employment_rate = 5L,
  CPI             = 34L,  # use 6L for four-period gross inflation
  CB_lending_rate = 50L,
  Emissions       = 7L,
  Energy_use      = 8L
)

parameter_names <- c(
  "Gamma", "chi", "psi1", "psi3", "alfa",
  "taylor1", "taylor2", "taylor", "kappa"
)

design <- read.csv(design_path, check.names = FALSE)

stopifnot(
  cop_id >= 1L,
  cop_id <= nrow(design),
  all(parameter_names %in% names(design))
)

cop <- design[cop_id, parameter_names, drop = FALSE]

dir.create(result_dir, recursive = TRUE, showWarnings = FALSE)
dir.create(
  file.path(result_dir, "errors"),
  recursive = TRUE,
  showWarnings = FALSE
)

# The model writes output beside its executable. Copy it to task-local scratch
# so parallel jobs do not overload or collide in one output directory.
scratch_root <- Sys.getenv("SLURM_TMPDIR", unset = tempdir())
job_dir <- tempfile(
  pattern = sprintf("dsk_cop%04d_", cop_id),
  tmpdir = scratch_root
)

dir.create(job_dir, recursive = TRUE)
on.exit(unlink(job_dir, recursive = TRUE, force = TRUE), add = TRUE)

local_exe <- file.path(job_dir, "dsk_SFC")
stopifnot(file.copy(source_exe, local_exe))
Sys.chmod(local_exe, mode = "0755")
stopifnot(file.access(local_exe, mode = 1) == 0L)

# Construct the input JSON for this CoP.
input <- fromJSON(base_json, simplifyVector = FALSE)

for (p in parameter_names) {
  input$params[[1]][[p]] <- as.numeric(cop[[p]])
}

input$params[[1]][["T"]] <- n_steps

json_path <- file.path(
  job_dir,
  sprintf("inputs_cop%04d.json", cop_id)
)

write_json(
  input,
  json_path,
  auto_unbox = TRUE,
  pretty = TRUE,
  digits = NA
)

result_array <- array(
  NA_real_,
  dim = c(length(output_columns), n_steps, n_mc),
  dimnames = list(
    variable = names(output_columns),
    step = as.character(seq_len(n_steps)),
    MC_run = as.character(seq_len(n_mc))
  )
)

attr(result_array, "CoP_ID") <- cop_id
attr(result_array, "parameters") <- unlist(cop, use.names = TRUE)
attr(result_array, "output_columns") <- output_columns
attr(result_array, "seed_scheme") <- "MC seed 1:n_mc, repeated across CoPs"

failures <- data.frame(
  seed = integer(),
  reason = character(),
  stringsAsFactors = FALSE
)

run_name <- sprintf("cop%04d", cop_id)

for (mc in seq_len(n_mc)) {
  # Reusing seeds 1:1000 for every CoP implements common random numbers,
  # which is normally useful when comparing parameter configurations.
  seed <- mc

  status <- suppressWarnings(
    system2(
      local_exe,
      args = c(
        shQuote(json_path),
        "-r", run_name,
        "-s", seed,
        "-f", 0,
        "-c", 0,
        "-v", 0
      ),
      stdout = FALSE,
      stderr = FALSE
    )
  )

  raw_file <- file.path(
    job_dir, "output",
    sprintf("results_%s_%d.txt", run_name, seed)
  )

  error_file <- file.path(
    job_dir, "output", "errors",
    sprintf("Errors_%s_%d.txt", run_name, seed)
  )

  error_nonempty <- (
    file.exists(error_file) &&
    !is.na(file.info(error_file)$size) &&
    file.info(error_file)$size > 0
  )

  values <- if (file.exists(raw_file)) {
    tryCatch(
      scan(raw_file, what = double(), quiet = TRUE),
      error = function(e) numeric()
    )
  } else {
    numeric()
  }

  valid_length <- length(values) == n_steps * n_model_columns

  if (status == 0L && !error_nonempty && valid_length) {
    model_output <- matrix(
      values,
      nrow = n_steps,
      ncol = n_model_columns,
      byrow = TRUE
    )

    selected <- t(
      model_output[, unname(output_columns), drop = FALSE]
    )

    if (all(is.finite(selected))) {
      result_array[, , mc] <- selected
    } else {
      failures <- rbind(
        failures,
        data.frame(seed = seed, reason = "Non-finite model output")
      )
    }
  } else {
    reason <- paste(
      if (status != 0L) sprintf("exit status %d", status),
      if (error_nonempty) "non-empty model error log",
      if (!valid_length) {
        sprintf(
          "expected %d values, obtained %d",
          n_steps * n_model_columns,
          length(values)
        )
      },
      collapse = "; "
    )

    failures <- rbind(
      failures,
      data.frame(seed = seed, reason = reason)
    )

    if (error_nonempty) {
      file.copy(
        error_file,
        file.path(
          result_dir, "errors",
          sprintf("cop%04d_seed%04d.txt", cop_id, seed)
        ),
        overwrite = TRUE
      )
    }
  }

  # Avoid retaining approximately 3 MB of raw 83-column output per run.
  unlink(c(raw_file, error_file), force = TRUE)

  if (mc %% 25L == 0L || mc == n_mc) {
    message(
      sprintf(
        "CoP %d: completed %d/%d runs; failures=%d",
        cop_id, mc, n_mc, nrow(failures)
      )
    )
  }
}

array_file <- file.path(
  result_dir,
  sprintf("cop_%04d.rds", cop_id)
)

saveRDS(result_array, array_file, compress = FALSE)

if (nrow(failures) > 0L) {
  write.csv(
    failures,
    file.path(result_dir, sprintf("cop_%04d_failures.csv", cop_id)),
    row.names = FALSE
  )
}

message("Saved ", array_file)

```

In your terminal, run a ten-seed pilot first:

Rscript run_one_cop.R \
  1 \
  calibration/Sample1.csv \
  Reissl_2025/dsk_sfc_inputs.json \
  Reissl_2025/build/dsk_SFC \
  simulation_results \
  10

Inspect it with:

```{r MC inspection}
x <- readRDS("simulation_results/cop_0001.rds")
dim(x)
# 8 600 10

attr(x, "parameters")
matplot(t(x["GDP", , ]), type = "l", lty = 1)
```

For the real run, replace 10 with 1000.

## Run the 1,000 CoPs on an HPC system

A Slurm array is the appropriate structure: one job per CoP, with 1,000 sequential seeds inside each job. In your terminal, run:

#!/bin/bash
#SBATCH --job-name=dsk_calibration
#SBATCH --array=1-1000%100
#SBATCH --cpus-per-task=1
#SBATCH --mem=1G
#SBATCH --time=12:00:00
#SBATCH --output=logs/cop_%A_%a.out
#SBATCH --error=logs/cop_%A_%a.err

cd "$SLURM_SUBMIT_DIR"

Rscript run_one_cop.R \
  "$SLURM_ARRAY_TASK_ID" \
  calibration/Sample1.csv \
  Reissl_2025/dsk_sfc_inputs.json \
  Reissl_2025/build/dsk_SFC \
  simulation_results \
  1000
  
Adjust %100 to the number of simultaneous jobs permitted by your cluster.

# Scale warning

This is a large experiment:

- Model executions: 1,000 × 1,000 = 1,000,000.
- One requested array: approximately 38.4 MB.
- All 1,000 arrays: approximately 38.4 GB before filesystem overhead.
- Unfiltered model output would total approximately 3 TB.

A baseline 600-step Release run took about 18 seconds in my local test, implying roughly 5,000 core-hours for the complete experiment. Run a pilot across several extreme CoPs before launching all one million simulations, use Release compilation, write transient files to node-local scratch, and delete each raw result immediately after extraction.
